--------------------------------------------------------------------------------
title: "Pharmacy First — Practice-level data- Medication Analysis"
author: "Arnaud Iradukunda"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
--------------------------------------------------------------------------------
# 1. Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(nlme)
library(geepack)
library(scales)
library(readr)
library(here)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# 2. Data and intervention dates

```{r data}
# I will read it using differnent path in OS
#practice_level_data <- read_csv(
#"C:/Users/arnau/Downloads/practice_level_data (8).csv"
#)

practice_level_data <- read_csv(
  here("output", "practice_level_data.csv.gz")
)

GP_connect <- as.Date("2025-10-01")
PF_start   <- as.Date("2024-02-01")

practice_level_data <- practice_level_data %>%
  clean_names() %>%  # This comes from the janitor package and standardises column names (no spaces).
  mutate(interval_start = as.Date(interval_start))

head(practice_level_data)
names(practice_level_data)

conditions <- c(
  "uti",
  "sinusitis",
  "insectbite",
  "otitismedia",
  "sorethroat",
  "shingles",
  "impetigo",
  "all_conditions"
)

control_map <- c(
  acutebronchitis_control         = "Acute bronchitis (control)",
  conjunctivitisallergic_control  = "Allergic conjunctivitis (control)",
  vulvovaginalcandidiasis_control = "Vulvovaginal candidiasis (control)"
)

head(practice_level_data)
#names(practice_level_data)
```

# 3. Monthly totals across practices
The dataset is aggregated to monthly totals across practices for GP and Pharmacy First (PF) medication counts.
```{r monthly-long}
monthly <- practice_level_data %>%
  group_by(interval_start) %>%
  summarise(
    population = sum(population, na.rm = TRUE),
    # Eligible populations
    #Conditions specific
    elig_uti         = sum(populationeligible_uuti, na.rm = TRUE),
    elig_sinusitis   = sum(populationeligible_sinusitis, na.rm = TRUE),
    elig_insectbite  = sum(populationeligible_insectbite, na.rm = TRUE),
    elig_otitismedia = sum(populationeligible_otitismedia, na.rm = TRUE),
    elig_sorethroat  = sum(populationeligible_sorethroat, na.rm = TRUE),
    elig_shingles    = sum(populationeligible_shingles, na.rm = TRUE),
    elig_impetigo    = sum(populationeligible_impetigo, na.rm = TRUE),
    # Controls specific
    elig_acutebronchitis_control = sum(populationeligible_acutebronchitis_control, na.rm = TRUE),
    elig_vulvovaginalcandidiasis_control= sum(populationeligible_vulvovaginalcandidiasis_control, na.rm = TRUE),
    elig_conjunctivitisallergic_control = sum(populationeligible_conjunctivitisallergic_control, na.rm = TRUE),
    across(
      starts_with("gp_medication_"),
      ~ sum(.x, na.rm = TRUE)   # Find all variables whose names start with gp_medication_ and sum each one within each month.
    ),
    across(
      starts_with("pf_medication_"),
      ~ sum(.x, na.rm = TRUE)  #  Find all variables whose names start with pf_medication_ and sum each one within each month (.x represents the current column being processed).
    ),
    .groups = "drop"          #  This remove the grouping (our dataset is now a normal data frame)
  ) %>%
  arrange(interval_start) %>%
  pivot_longer(                                               
    cols = starts_with(c("gp_medication_", "pf_medication_")), # GP/PF medication (wide to long)
    names_to = c("Setting", "condition"),                      # Setting = gp or Setting = pf, condition = uti

    names_pattern = "^(gp|pf)_medication_(.*)$",             #
    values_to = "medication_count"
  ) %>%
  mutate(
    Setting = toupper(Setting),
    time_index = match(
      interval_start,
      sort(unique(interval_start))
    ) - 1,
    post = as.integer(interval_start >= GP_connect),
    time_since_int = if_else(
      post == 1,
      time_index - min(time_index[post == 1]),
      0
    )
  ) %>%
  arrange(interval_start, Setting, condition)
names(monthly)
```
# Monthly_long ( this pivot eligible pop into long-format)
```{r monthly_long}
library(dplyr)
library(tidyr)
monthly_long <- monthly %>%
  pivot_longer(
    cols = starts_with("elig_"),
    names_to = "eligible_condition",
    values_to = "eligible_population"
  ) %>%
  mutate(
    condition = case_when(
      eligible_condition == "elig_uti" ~ "uti",
      eligible_condition == "elig_sinusitis" ~ "sinusitis",
      eligible_condition == "elig_insectbite" ~ "insectbite",
      eligible_condition == "elig_otitismedia" ~ "otitismedia",
      eligible_condition == "elig_sorethroat" ~ "sorethroat",
      eligible_condition == "elig_shingles" ~ "shingles",
      eligible_condition == "elig_impetigo" ~ "impetigo",
      eligible_condition == "elig_acutebronchitis_control" ~ "acutebronchitis_control",
      eligible_condition == "elig_vulvovaginalcandidiasis_control" ~ "vulvovaginalcandidiasis_control",
      eligible_condition == "elig_conjunctivitisallergic_control" ~ "conjunctivitisallergic_control"
    )
  ) %>%
  select(
    interval_start,
    population,
    Setting,
    condition,
    eligible_population,
    medication_count,
    time_index,
    post,
    time_since_int
  )
names(monthly_long)
table(monthly_long$condition)
table(monthly_long$eligible_population)
```

## 6. Descriptive analysis
## 6.1 Medication numerator over time

```{r}
table(monthly_long$condition)
```


```{r descriptive-numerator}
descr_numerator <- monthly_long %>%            # Here i can also use monthly long
  filter(condition %in% conditions) %>%
  select(
    interval_start,
    Setting,
    condition,
    medication_count
  ) %>%
  mutate(
    condition = str_to_sentence(condition)
  )
```
#

```{r}
table(descr_numerator$Setting)
```
# Graph 1: Medication counts
```{r graph_A1}
graph_A1_numerator <- ggplot(
  descr_numerator,
  aes(
    x = interval_start,
    y = medication_count,
    colour = Setting
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dashed",
    colour = "black"
  ) +
  facet_wrap(
    ~ condition,
    scales = "free_y",
    ncol = 1
  ) +
  scale_x_date(
    date_labels = "%b %y",
    date_breaks = "1 month"
  ) +
  scale_colour_manual(
    values = c(
      "GP" = "steelblue",
      "PF" = "firebrick"
    )
  ) +
  labs(
    title = "Medication counts by condition and setting",
    subtitle = "Dashed line = GP Connect",
    x = NULL,
    y = "Number of medications (numerator)",
    colour = "Setting"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
graph_A1_numerator
```

## 6.2 Medication rate over time

```{r descriptive-num}
descr_rate <- monthly_long %>%
  filter(condition %in% conditions) %>%
  select(
    interval_start,
    Setting,
    condition,
    eligible_population,
    medication_count
  ) %>%
  mutate(
    condition = str_to_sentence(condition), # This capitalises the first letter and makes the rest lowercase.
    rate_per_1000 = (medication_count/eligible_population)*1000
  )
```

#graph 2: Medication rate
```{r graph_A2_rate}
graph_A2_rate <- ggplot(
  descr_rate,
  aes(
    x = interval_start,
    y = rate_per_1000,
    colour = Setting
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dashed",
    colour = "black"
  ) +
  facet_wrap(
    ~ condition,
    scales = "free_y",
    ncol = 1
  ) +
  scale_x_date(
    date_labels = "%b %y",
    date_breaks = "1 month"
  ) +
  scale_colour_manual(
    values = c(
      "GP" = "steelblue",
      "PF" = "darkred"
    )
  ) +
  labs(
    title = "Medication rate by condition and setting",
    subtitle = "Rate per 1,000 population; dashed line = GP Connect",
    x = NULL,
    y = "Rate per 1,000 population",
    colour = "Setting"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
graph_A2_rate
```
# 7. GP intervention-control descriptive plots
```{r intervention-control-data}
descr_pairs <- practice_level_data %>%
  select(
    interval_start,
    gp_medication_uti,
    gp_medication_vulvovaginalcandidiasis_control,
    gp_medication_sorethroat,
    gp_medication_acutebronchitis_control,
    gp_medication_insectbite,
    gp_medication_conjunctivitisallergic_control
  ) %>%
  group_by(interval_start) %>%
  mutate(
    medication_uti =
      sum(gp_medication_uti, na.rm = TRUE),

    medication_vulvovaginalcandidiasis_control =
      sum(gp_medication_vulvovaginalcandidiasis_control, na.rm = TRUE),

    medication_sorethroat =
      sum(gp_medication_sorethroat, na.rm = TRUE),

    medication_acutebronchitis_control =
      sum(gp_medication_acutebronchitis_control, na.rm = TRUE),

    medication_insectbite =
      sum(gp_medication_insectbite, na.rm = TRUE),

    medication_conjunctivitisallergic_control =
      sum(gp_medication_conjunctivitisallergic_control, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  select(
    interval_start,
    starts_with("medication_")
  ) %>%
  distinct() %>%
  pivot_longer(
    cols = starts_with("medication_"),
    names_to = "condition",
    values_to = "count"
  ) %>%
  mutate(
    condition = recode(
      condition,
      medication_uti =
        "UTI",
      medication_vulvovaginalcandidiasis_control =
        "Vulvovaginal candidiasis",
      medication_sorethroat =
        "Sore throat",
      medication_acutebronchitis_control =
        "Acute bronchitis",
      medication_insectbite =
        "Insect bite",
      medication_conjunctivitisallergic_control =
        "Allergic conjunctivitis"
    )
  )
```

## 7.1 UTI vs vulvovaginal candidiasis
```{r descr_pairs}
table(descr_pairs$condition)
```
## Graph
```{r graph-uti-and-control}
graph_A3_uti <- descr_pairs %>%
  filter(condition %in%c("UTI","Vulvovaginal candidiasis")) %>%
  ggplot(
    aes(
      x = interval_start,
      y = count,
      colour = condition
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(
    date_labels = "%b %y",
    date_breaks = "1 month"
  ) +
  labs(
    title = "GP medications for UTI and Vulvovaginal candidiasis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Number of medications (numerator)",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )
  
graph_A3_uti
```
## 7.2 Sore throat vs acute bronchitis

```{r graph-sorethroat-control}
graph_A3_sorethroat <- descr_pairs %>%
  filter(
    condition %in%
      c(
        "Sore throat",
        "Acute bronchitis"
      )
  ) %>%
  ggplot(
    aes(
      x = interval_start,
      y = count,
      colour = condition
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(
    date_labels = "%b %y",
    date_breaks = "1 month"
  ) +
  labs(
    title = "GP medications for sore throat and acute bronchitis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Number of medications (numerator)",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )
graph_A3_sorethroat
```

## 7.3 Insect bite vs allergic conjunctivitis

```{r graph-insectbite-control}
graph_A3_insectbite <- descr_pairs %>%
  filter(
    condition %in%
      c(
        "Insect bite",
        "Allergic conjunctivitis"
      )
  ) %>%
  ggplot(
    aes(
      x = interval_start,
      y = count,
      colour = condition
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(
    date_labels = "%b %y",
    date_breaks = "1 month"
  ) +
  labs(
    title = "GP medications for insect bite and allergic conjunctivitis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Number of medications (numerator)",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_A3_insectbite
```



#Pair rate

#Descriptives pairs rate
```{r descr_pairs_rate}
descr_pairs_rate <- monthly_long %>%
  select(
    interval_start,
    Setting,
    condition,
    eligible_population,
    medication_count
  ) %>%
  mutate(
    condition = str_to_sentence(condition), # This capitalises the first letter and makes the rest lowercase.
    rate_per_1000 = (medication_count/eligible_population)*1000
  )
```

## 8.1 UTI vs vulvovaginal candidiasis rate
```{r}
table(descr_pairs_rate$condition)
```


```{r graph-uti-control}
graph_B3_uti <- descr_pairs_rate %>%
  filter(
    condition %in%
      c(
        "Uti",
        "Conjunctivitisallergic_control"
      )
  ) %>%
  ggplot(
    aes(
      x = interval_start,
      y = rate_per_1000,
      colour = condition
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(
    date_labels = "%b %y",
    date_breaks = "1 month"
  ) +
  labs(
    title = "GP medications for UTI and Vulvovaginal candidiasis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Rate per 1,000 population",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_B3_uti
```

## 8.2 Sore throat vs acute bronchitis rate
```{r graph-sorethroat-control}
graph_B3_sorethroat <- descr_pairs_rate %>%
  filter(
    condition %in%
      c(
        "Sorethroat",
        "Acutebronchitis_control"
      )
  ) %>%
  ggplot(
    aes(
      x = interval_start,
      y = rate_per_1000,
      colour = condition
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(
    date_labels = "%b %y",
    date_breaks = "1 month"
  ) +
  labs(
    title = "GP medications for sore throat and acute bronchitis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Rate per 1000 population",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_B3_sorethroat
```

## 8.3 Insect bite vs allergic conjunctivitis rate
```{r graph-insectbite-control}
graph_B3_insectbite <- descr_pairs_rate %>%
  filter(
    condition %in%
      c(
        "Insectbite",
        "Conjunctivitisallergic_control"
      )
  ) %>%
  ggplot(
    aes(
      x = interval_start,
      y = rate_per_1000,
      colour = condition
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(
    date_labels = "%b %y",
    date_breaks = "1 month"
  ) +
  labs(
    title = "GP medications for insect bite and allergic conjunctivitis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Rate per 1,000 population",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_B3_insectbite
```

# 8. ITSA models — GP
The following section prepares an interrupted time-series analysis for all seven Pharmacy First conditions.
The intervention is **PF service implementation on 1 february 2024**.
However,**GP Connect on 1 October 2025** may influence reported data in GP records.

We model the monthly GP-recorded medication counts using a segmented
Poisson regression with an offset for the population at risk.

The model specification is:

\[
\log\left(E(Y_{gt})\right)
=
\beta_0
+
\beta_1 T_{gt}
+
\beta_2 X_{gt}
+
\beta_3 X_{gt}T_{gt}
+
\log(Pop_{gt})
\tag{7}
\]

where:

- \(Y_{gt}\) is the observed count of the outcome for practice \(g\) at month \(t\);
- \(E(Y_{gt})\) is the expected count;
- \(T_{gt}\) is the time in months from the start of the study period;
- \(X_{gt}\) is an indicator for the post-intervention period
  (\(1\) = at/after PF service implementation, \(0\) = before);
- \(X_{gt}T_{gt}\) is the interaction term allowing the slope to change after the intervention;
- \(Pop_{gt}\) is the population at risk for practice \(g\) at month \(t\), included as an offset.

- `β0` = baseline level of the outcome at the start of the study period;
- `β1` = baseline time trend (change in the outcome per month before PF service);
- `β2` = immediate level change in the outcome following PF service implementation;
- `β3` = change in the time trend (slope) following PF service implementation.


## 8.1 Prepare ITSA dataset

```{r}
#names(practice_level_data)
# GP Connect intervention date
intervention_date <- as.Date("2025-10-01")
#ITSA dataset

practice_level_data_itsa <- practice_level_data %>%
  mutate( interval_start = as.Date(interval_start)) %>%
  arrange(practice,interval_start) %>%
  group_by(practice) %>%
  mutate(
    # Time starts at 1 for each practice
    T = row_number()-1,
    # Intervention start
    X = as.integer(interval_start >= intervention_date),
    # Time point of intervention
    first_post_T = T[
      interval_start == intervention_date
    ][1],

    # Change in slope after  intervention
    XT = pmax(
      T - first_post_T,
      0
    ) * X
  ) %>%
  ungroup() %>%
  select(-first_post_T)

# Check data
print(head(practice_level_data_itsa, 14))

# Check ITSA variables
print(
  practice_level_data_itsa %>%
    distinct(interval_start, T, X, XT) %>%
    arrange(interval_start)
)
# Export into csv
#write_csv(practice_level_data_itsa,"practice_level_data_itsa.csv")
write_csv(
  practice_level_data_itsa,
  "practice_level_data_itsa.csv"
)
```




























## 8.1 Prepare ITSA dataset
# Eligible population
```{r}
names(practice_level_data)[
  names(practice_level_data) == "populationeligible_uuti"
] <- "populationeligible_uti"

itsa_gp_eligible <- practice_level_data %>%
  select(
    practice,
    interval_start,
    starts_with("populationeligible_")
  )%>%
  distinct()

# Eligible population to long format
itsa_gp_eligible_long <- itsa_gp_eligible %>%
  pivot_longer(
    cols = starts_with("populationeligible_"),
    names_to = "condition",
    names_prefix = "populationeligible_",
    values_to = "eligible_population"
  )
# GP medication
itsa_gp_medication <- practice_level_data %>%
  select(
    practice,
    interval_start,
    starts_with("gp_medication_")
  )%>%
  distinct()
# GP medication to long format
itsa_gp_medication_long <- itsa_gp_medication %>%
  pivot_longer(
    cols = starts_with("gp_medication_"),
    names_to = "condition",
    names_prefix = "gp_medication_",
    values_to = "medication_count"
  )
# Both (eligible population and GP medication)
itsa_gp_data_long <- itsa_gp_eligible_long %>%
  left_join(
    itsa_gp_medication_long,
    by = c("practice", "interval_start", "condition")
  )
```
## II.GEE models:
## II.1.GP UTI medication prescribing
```{r gee_uti_data}

# 1. Load packages
library(dplyr)
library(ggplot2)
library(lubridate)
library(geepack)
library(broom)

#2.  Data preparation
gee_data <- practice_level_data %>%
  clean_names() %>%
  mutate(
    interval_start = as.Date(interval_start),

    # Time in months from February 2022
    # Feb 2022 = 0
    time = (year(interval_start) - 2024) * 12 +
           month(interval_start) - 6,

    # Pharmacy First implementation:
    # 1 February 2024
    post_pf = as.integer(
      interval_start >= as.Date("2024-10-01")
    ),

    # Months since Pharmacy First implementation
    time_after_pf = if_else(
      post_pf == 1,
      time - 4,
      0
    )
  )
```
#gee_uti_data
```{r gee_uti_data}
gee_uti_data <- gee_data %>%
  select(
    practice,
    interval_start,
    populationeligible_uti,
    gp_medication_uti,
    time,
    post_pf,
    time_after_pf
  )
```

#3. Data checks
```{r data-check}

glimpse(gee_uti_data)
summary(gee_uti_data$gp_medication_uti)
summary(gee_uti_data$populationeligible_uti)
table(gee_uti_data$post_pf)
range(gee_uti_data$interval_start, na.rm = TRUE)
```
# 4.Missing and invalid values
```{r missingness}
 gee_uti_data %>%
  summarise(
    n_rows = n(),
    n_practices = n_distinct(practice),
    missing_outcome = sum(is.na(gp_medication_uti)),
    missing_population = sum(is.na(population)),
    zero_population = sum(population <= 0, na.rm = TRUE),
    missing_date = sum(is.na(interval_start))
  )
```
# 5. Fit the GEE model
```{r gee_uti_model}
gee_uti_model <- geeglm(
  gp_medication_uti ~ time + post_pf + time_after_pf,
  family = poisson(link = "log"),
  offset = log(population),
  id = practice,
  corstr = "ar1",
  data = gee_uti_data
)
```
# 6. Model summary
```{r model_summary}
summary(gee_uti_model)
```

# 7. Incidence-rate ratios (IRRs)
```{r irr}
irr_results <- broom::tidy(
  gee_uti_model,
  exponentiate = TRUE,
  conf.int = TRUE
)

irr_results
```
# 8. Extract the main Pharmacy First effects
```{r pf_effects}
pf_effects <- irr_results %>%
  filter(term %in% c("post_pf", "time_after_pf")) %>%
  select(
    term,
    estimate,
    conf.low,
    conf.high,
    p.value
  )

pf_effects
```
# 9.Table
```{r pf_effects_table}
pf_effects %>%
  mutate(
    IRR = round(estimate, 3),
    CI = paste0(
      round(conf.low, 3),
      "–",
      round(conf.high, 3)
    ),
    p_value = signif(p.value, 3)
  ) %>%
  select(
    term,
    IRR,
    CI,
    p_value
  )
```
# 10. Observed monthly GP UTI medication rate
```{r observed_rate}
monthly_uti <- gee_uti %>%
  group_by(interval_start) %>%
  summarise(
    gp_medication_uti = sum(
      gp_medication_uti,
      na.rm = TRUE
    ),

    population = sum(
      population,
      na.rm = TRUE
    ),

    rate_per_1000 = (
      gp_medication_uti / population
    ) * 1000,

    .groups = "drop"
  )
```
# 11. Visualise observed prescribing rate
```{r observed_rate_plot}
ggplot(
  monthly_uti,
  aes(
    x = interval_start,
    y = rate_per_1000
  )
) +

  geom_line(linewidth = 0.8) +

  geom_point(size = 1.5) +

  # Pharmacy First implementation
  geom_vline(
    xintercept = as.numeric(
      as.Date("2024-02-01")
    ),
    linetype = "dashed",
    linewidth = 0.8
  ) +

  annotate(
    "text",
    x = as.Date("2024-02-01"),
    y = Inf,
    label = "Pharmacy First",
    vjust = 1.5,
    hjust = -0.05
  ) +

  labs(
    title = "GP UTI antimicrobial prescribing",
    subtitle = "Monthly prescribing rate before and after Pharmacy First implementation",
    x = "Month",
    y = "GP UTI medication prescriptions per 1,000 population"
  ) +

  theme_minimal() +

  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```
#
# 12.Fitted/predicted trend from the GEE model
```{r predicted_trend}
gee_uti <- gee_uti %>%
  mutate(
    predicted_rate_per_1000 =
      predict(
        gee_uti_model,
        type = "response"
      ) /
      population * 1000
  )
```
# 13. Plot: observed and model-predicted rates
```{r observed_and_predicted_plot}
predicted_monthly <- gee_uti %>%
  group_by(interval_start) %>%
  summarise(
    observed_rate =
      sum(gp_medication_uti, na.rm = TRUE) /
      sum(population, na.rm = TRUE) * 1000,

    predicted_rate =
      sum(
        predicted_rate_per_1000 * population,
        na.rm = TRUE
      ) /
      sum(population, na.rm = TRUE),

    .groups = "drop"
  )


ggplot(
  predicted_monthly,
  aes(x = interval_start)
) +

  geom_line(
    aes(y = observed_rate),
    linewidth = 0.8
  ) +

  geom_line(
    aes(y = predicted_rate),
    linewidth = 1,
    linetype = "dashed"
  ) +

  geom_vline(
    xintercept = as.numeric(
      as.Date("2024-02-01")
    ),
    linetype = "dotted",
    linewidth = 0.8
  ) +

  annotate(
    "text",
    x = as.Date("2024-02-01"),
    y = Inf,
    label = "Pharmacy First",
    vjust = 1.5,
    hjust = -0.05
  ) +

  labs(
    title = "GP UTI antimicrobial prescribing",
    subtitle = "Observed and GEE-modelled monthly prescribing rates",
    x = "Month",
    y = "Prescriptions per 1,000 population"
  ) +

  theme_minimal() +

  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
# 14.Intervention segments
```{r intervention_segments}
plot<-ggplot(
  monthly_uti,
  aes(
    x = interval_start,
    y = rate_per_1000
  )
) +

  geom_line(linewidth = 0.8) +

  geom_point(size = 1.3) +

  geom_vline(
    xintercept = as.numeric(
      as.Date("2024-02-01")
    ),
    linetype = "dashed",
    linewidth = 0.8
  ) +

  annotate(
    "text",
    x = as.Date("2024-02-01"),
    y = max(
      monthly_uti$rate_per_1000,
      na.rm = TRUE
    ),
    label = "Pharmacy First\n1 February 2024",
    hjust = -0.05,
    vjust = 1
  ) +

  labs(
    title = "Interrupted time series of GP UTI prescribing",
    x = "Month",
    y = "GP UTI medication rate per 1,000 population"
  ) +

  theme_minimal()
plot
  
```

### Interpretation 

The model is:

$$
\log(\mu_{it}) =
\beta_0
+\beta_1Time_t
+\beta_2PostPF_t
+\beta_3TimeAfterPF_t
+\log(Population_{it})
$$

where:

* **β₀** = baseline GP UTI medication prescribing rate
* **β₁** = pre-Pharmacy First monthly trend
* **β₂** = **immediate level change** after Pharmacy First
* **β₃** = **change in slope** after Pharmacy First

Because this is a Poisson model with a log link, the exponentiated coefficients are **rate ratios**.

For example:

```{r }
exp(coef(gee_uti_model)["post_pf"])
```

gives the immediate rate ratio.

And:

```{r }
exp(coef(gee_uti_model)["time_after_pf"])
```

gives the multiplicative change in the prescribing rate for each additional month after Pharmacy First.



