---
title: "WP2P4-Descriptive Analysis: summary tables and figures"
author: "Arnaud"
date: "2026-07-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(readr)
library(knitr)
library(here)
library(tidyverse)
library(lubridate)
library(scales)
library(dplyr)
```

# Dataset

```{r}
dataset <- read_csv(
  here("output", "dataset_patients_combined.csv.gz"),
  col_types = cols(
    patient_id = col_integer(),
    age = col_double()
  )
)

dataset <- dataset %>%
  select(patient_id,index_date,sex,age,imd,ethnicity,stp,region,has_medication
)
dataset$index_date <- as.Date(dataset$index_date)
head(dataset)
```
# Table 1. Distribution of patients receiving medication by IMD
```{r}
table_imd <- dataset %>%
  group_by(imd) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(imd)

kable(
  table_imd,
  caption = "Table 1. Distribution of patients receiving medication by IMD."
)
```

# Figure 1. Patients receiving medication over time by IMD

```{r fig.width=10, fig.height=5, fig.cap="Figure 1. Patients receiving medication over time by IMD."}
table_imd_plot <- dataset %>%
  group_by(index_date, imd) %>%
  summarise(
    n_patients = sum(has_medication == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(imd != "Missing")

ggplot(
  table_imd_plot,
  aes(
    x = index_date,
    y = n_patients,
    colour = imd,
    group = imd
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "IMD"
  ) +
  scale_x_date(
    date_breaks = "3 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

# Table 2. Distribution of patients receiving medication by Region

```{r}
table_region <- dataset %>%
  group_by(region) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(region)

kable(
  table_region,
  caption = "Table 2. Distribution of patients receiving medication by Region."
)
```

# Figure 2. Patients receiving medication over time by Region

```{r fig.width=10, fig.height=5, fig.cap="Figure 2. Patients receiving medication over time by Region."}
table_region_plot <- dataset %>%
  group_by(index_date, region) %>%
  summarise(
     patients = sum(has_medication == 1, na.rm = TRUE), #n_patients = n(),
    .groups = "drop"
  ) %>%
  filter(region != "Missing")

ggplot(
  table_region_plot,
  aes(
    x = index_date,
    y = patients,
    colour = region,
    group = region
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "Region"
  ) +
  scale_x_date(
    date_breaks = "3 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```
# Table 3. Distribution of patients receiving medication by Sex

```{r}
table_sex <- dataset %>%
  group_by(sex) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(sex)

kable(
  table_sex,
  caption = "Table 3. Distribution of patients receiving medication by sex."
)
```

# Figure 3. Patients receiving medication over time by Sex

```{r fig.width=10, fig.height=5, fig.cap="Figure 3. Patients receiving medication over time by sex."}
table_sex_plot <- dataset %>%
  group_by(index_date, sex) %>%
  summarise(
     patients = sum(has_medication == 1, na.rm = TRUE), #n_patients = n(),
    .groups = "drop"
  ) %>%
  filter(!is.na(sex))

ggplot(
  table_sex_plot,
  aes(
    x = index_date,
    y = patients,
    colour = sex,
    group = sex
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "Sex"
  ) +
  scale_x_date(
    date_breaks = "3 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

# Table 4. Distribution of patients receiving medication by Ethnicity

```{r}
table_ethnicity <- dataset %>%
  group_by(ethnicity) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(ethnicity)

kable(
  table_ethnicity,
  caption = "Table 4. Distribution of patients receiving medication by ethnicity."
)
```

# Figure 4. Patients receiving medication over time by Ethnicity

```{r fig.width=10, fig.height=5, fig.cap="Figure 4. Patients receiving medication over time by ethnicity."}
table_ethnicity_plot <- dataset %>%
  group_by(index_date, ethnicity) %>%
  summarise(
    n_patients = sum(has_medication == 1, na.rm = TRUE), #n_patients = n(),
    .groups = "drop"
  ) %>%
  filter(
    !is.na(ethnicity),
    ethnicity != "Missing"
  )

ggplot(
  table_ethnicity_plot,
  aes(
    x = index_date,
    y = n_patients,
    colour = ethnicity,
    group = ethnicity
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "Ethnicity"
  ) +
  scale_x_date(
    date_breaks = "3 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```
