---
title: "WP2P4-Descriptive Analysis: summary tables and figures"
author: "Arnaud"
date: "2026-07-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(readr)
library(knitr)
library(here)
library(tidyverse)
library(lubridate)
library(scales)
library(dplyr)
```

# Dataset

```{r}
dataset <- read_csv(
  here("output", "dataset_patients_combined.csv.gz"),
  col_types = cols(
    patient_id = col_integer(),
    age = col_double()
  )
)

dataset <- dataset %>%
  select( 
  patient_id,
  index_date,sex,
  imd,ethnicity,
  stp,region,
  has_medication,numerator_gp_nitrofurantoin_uti,
  numerator_pf_nitrofurantoin_uti
)
dataset$index_date <- as.Date(dataset$index_date)
head(dataset)
```

# Table 0_0. Nitrofurantoin prescribing volume by region
```{r}
table_0_0 <- dataset %>%
  group_by(region) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),

    gp_nitro_volume = sum(numerator_gp_nitrofurantoin_uti, na.rm = TRUE),
    pf_nitro_volume = sum(numerator_pf_nitrofurantoin_uti, na.rm = TRUE),
    total_nitro_volume = gp_nitro_volume + pf_nitro_volume,

    .groups = "drop"
  ) %>%
  arrange(region)

kable(
  table_0_0,
  caption = "Table 0_0. Nitrofurantoin prescribing volume by region."
)
```
#
```{r fig.width=10, fig.height=5, fig.cap="Figure 0_0. Nitrofurantoin prescribing volume over time by region."}
table_region_plot <- dataset %>%
  group_by(index_date, region) %>%
  summarise(
    gp_nitro_volume = sum(numerator_gp_nitrofurantoin_uti, na.rm = TRUE),
    pf_nitro_volume = sum(numerator_pf_nitrofurantoin_uti, na.rm = TRUE),
    total_nitro_volume = gp_nitro_volume + pf_nitro_volume,
    .groups = "drop"
  ) %>%
  filter(region != "Missing")

nitro_region_plot <- ggplot(
  table_region_plot,
  aes(
    x = index_date,
    y = total_nitro_volume,
    colour = region,
    group = region
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Nitrofurantoin prescribing volume",
    colour = "Region"
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

nitro_region_plot

# Save as PNG
ggsave(
  filename = "results_0_0_region_nitrofurantoin_volume.png",
  plot = nitro_region_plot,
  path = here::here("results_Arnaud"),
  width = 10,
  height = 5,
  dpi = 300
)
```
# Table 0_1. Distribution of patients receiving medication by STP
```{r}
table_stp <- dataset %>%
  group_by(stp) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(stp)

kable(
  table_stp,
  caption = "Table 0_1. Distribution of patients receiving medication by STP."
)
```

# Table 1. Distribution of patients receiving medication by IMD
```{r}
table_imd <- dataset %>%
  group_by(imd) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(imd)

kable(
  table_imd,
  caption = "Table 1. Distribution of patients receiving medication by IMD."
)
```

# Figure 1. Patients receiving medication over time by IMD 

```{r fig.width=10, fig.height=5, fig.cap="Figure 1. Patients receiving medication over time by IMD."}
table_imd_plot <- dataset %>%
  group_by(index_date, imd) %>%
  summarise(
    n_patients = sum(has_medication == 0, na.rm = TRUE), # We will use 1 when using real data
    .groups = "drop"
  ) %>%
  filter(imd != "Missing")

imd_plot <- ggplot(
  table_imd_plot,
  aes(
    x = index_date,
    y = n_patients,
    colour = imd,
    shape = imd,
    group = imd
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "IMD",
    shape = "IMD"
    
  ) +
  scale_x_date(
    date_breaks = "1 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
  imd_plot 
 # Save as PNG 
ggsave(
  filename = "results_0_imd_plot.png",
  plot = imd_plot,
  path = here::here("results_Arnaud"),
  width = 7,
  height = 4,
  dpi = 300
)
```

# Table 2. Distribution of patients receiving medication by Region

```{r}
table_region <- dataset %>%
  group_by(region) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(region)

kable(
  table_region,
  caption = "Table 2. Distribution of patients receiving medication by Region."
)
```

# Figure 2. Patients receiving medication over time by Region

```{r fig.width=10, fig.height=5, fig.cap="Figure 2. Patients receiving medication over time by Region."}
table_region_plot <- dataset %>%
  group_by(index_date, region) %>%
  summarise(
     patients = sum(has_medication == 0, na.rm = TRUE), #n_patients = n(),
    .groups = "drop"
  ) %>%
  filter(region != "Missing")

region_plot <- ggplot(
  table_region_plot,
  aes(
    x = index_date,
    y = patients,
    colour = region,
    group = region
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "Region"
  ) +
  scale_x_date(
    date_breaks = "1 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
ggsave(
  filename = "results_0_region_plot.png",
  plot = region_plot,
  path = here::here("results_Arnaud"),
  width = 7,
  height = 4,
  dpi = 300
)
```
# Table 3. Distribution of patients receiving medication by Sex

```{r}
table_sex <- dataset %>%
  group_by(sex) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(sex)

kable(
  table_sex,
  caption = "Table 3. Distribution of patients receiving medication by sex."
)
```

# Figure 3. Patients receiving medication over time by Sex

```{r fig.width=10, fig.height=5, fig.cap="Figure 3. Patients receiving medication over time by sex."}

table_sex_plot <- dataset %>%
  group_by(index_date, sex) %>%
  summarise(
    patients = sum(has_medication == 0, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(sex))

sex_plot <- ggplot(
  table_sex_plot,
  aes(
    x = index_date,
    y = patients,
    colour = sex,
    group = sex
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "Sex"
  ) +
  scale_x_date(
    date_breaks = "1 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

sex_plot

ggsave(
  filename = "results_0_sex_plot.png",
  plot = sex_plot,
  path = here::here("results_Arnaud"),
  width = 7,
  height = 4,
  dpi = 300
)
```

# Table 4. Distribution of patients receiving medication by Ethnicity

```{r}
table_ethnicity <- dataset %>%
  group_by(ethnicity) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(ethnicity)

kable(
  table_ethnicity,
  caption = "Table 4. Distribution of patients receiving medication by ethnicity."
)
```

# Figure 4. Patients receiving medication over time by Ethnicity

```{r fig.width=10, fig.height=5, fig.cap="Figure 4. Patients receiving medication over time by ethnicity."}
table_ethnicity_plot <- dataset %>%
  group_by(index_date, ethnicity) %>%
  summarise(
    n_patients = sum(has_medication == 0, na.rm = TRUE), #n_patients = n(),
    .groups = "drop"
  ) %>%
  filter(
    !is.na(ethnicity),
    ethnicity != "Missing"
  )

ethnicity_plot <- ggplot(
  table_ethnicity_plot,
  aes(
    x = index_date,
    y = n_patients,
    colour = ethnicity,
    group = ethnicity
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "Ethnicity"
  ) +
  scale_x_date(
    date_breaks = "1 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
ggsave(
  filename = "results_0_ethnicity_plot.png",
  plot = ethnicity_plot,
  path = here::here("results_Arnaud"),
  width = 7,
  height = 4,
  dpi = 300
)
```

# Model 

```{r}
#Measures
library(readr)
#measures <- read_csv("C:/Users/arnau/Downloads/data_patients_measures_Arnaud (2).csv")

measures <- read_csv(
  here("output", "data_patients_measures_Arnaud.csv")
names(measures)
table(measures$measure)
measures<-measures|>
  filter(measure == "gp_medication_all_conditions")
names(measures)
head(measures)
table(measures$measure)
# covariates.
# T : months since start of study :Feb 2022
# X : Intervention starts 1 Feb 2024
# XT: Interaction term (Change in slope)
library(dplyr)
library(lubridate)
measures <- measures %>%
  mutate(
    month_date = as.Date(interval_start),
    
    # T: months since start of study
    T = row_number() - 1
  ) %>%
  group_by(practice) %>%
  arrange(month_date, .by_group = TRUE) %>%
  mutate(
    T = row_number() - 1
  ) %>%
  ungroup() %>%
  mutate(
    
    # Intervention starts 1 Feb 2024
    X = if_else(month_date >= as.Date("2024-02-01"), 1, 0),
    
    # Change in slope
    XT = if_else(
      X == 1,
      T - min(T[X == 1]),
      0
    )
  )
head(measures)
library(geepack)
gee_model_1 <- geeglm(
  numerator ~ T + X + XT +
    offset(log(denominator)),
  id = practice,
  data = measures,
  family = poisson(link = "log"),
  corstr = "ar1"
)
summary(gee_model_1)
#2.simulated numerator and denominator
set.seed(12345) 
measures <- measures %>%
  mutate(
    
    # Simulate denominator (consultations)
    denominator = round(
      rnorm(
        n(),
        mean = 120,
        sd = 20
      )
    ),
    
    denominator = pmax(denominator, 30),
    
    # Baseline prescribing probability (~65%)
    prescribing_prob = 0.65 -
      
      # Gradual decline over study
      0.0015 * T -
      
      # Immediate reduction after Pharmacy First
      0.05 * X -
      
      # Continued decline after intervention
      0.001 * XT,
    
    prescribing_prob = pmin(
      pmax(prescribing_prob, 0.10),
      0.95
    ),
    # Simulated numerator
    numerator = rbinom(
      n(),
      size = denominator,
      prob = prescribing_prob
    ),
    # Prescribing rate
    prescribing_rate = numerator / denominator
  )
head(measures)
#
library(geepack)
gee_model_2 <- geeglm(
numerator ~ T + X + XT +
  offset(log(denominator)),
id = practice,
data = measures,
family = poisson(link = "log"),
corstr = "ar1"
)
summary(gee_model_2)

#Predictions :
measures$pred <- predict(
gee_model,
type = "response"
)
#Aggregation(for plotting)

library(dplyr)
plot_data <- measures %>%
  group_by(month_date) %>%
  summarise(
    observed = mean(numerator, na.rm = TRUE),
    predicted = mean(pred, na.rm = TRUE),
    .groups = "drop"
  )
#ITS plot
ggplot(plot_data, aes(x = month_date)) +
  geom_line(aes(y = observed, colour = "Observed"), linewidth = 1.1) +
  geom_line(aes(y = predicted, colour = "Predicted"), linewidth = 1.1) +
  geom_vline(
    xintercept = as.Date("2024-02-01"),
    linetype = "dashed",
    colour = "red",
    linewidth = 1
  ) +
  scale_colour_manual(values = c(
    Observed = "black",
    Predicted = "blue"
  )) +
  
  labs(
    x = "Month",
    y = "GP medication count",
    colour = "",
    title = "Interrupted Time Series using GEE",
    subtitle = "Outcome: GP medication prescriptions for all Pharmacy First conditions"
  ) +
  
  theme_bw(base_size = 14)

#------------------------------------------
# Predicted values from the fitted model
data$fitted <- predict(gee_model, type = "response")

# Counterfactual data: assume no intervention occurred
cf_data <- data

cf_data$X <- 0
cf_data$XT <- 0

# Counterfactual prediction
cf_data$counterfactual <- predict(
  gee_model,
  newdata = cf_data,
  type = "response"
)
data$counterfactual <- cf_data$counterfactual


#
library(dplyr)
#Aggregate to monthly averages
plot_data <- data %>%
  group_by(month_date) %>%
  summarise(
    observed = mean(numerator, na.rm = TRUE),
    fitted = mean(fitted, na.rm = TRUE),
    counterfactual = mean(counterfactual, na.rm = TRUE),
    .groups = "drop"
  )
#
library(ggplot2)
ggplot(plot_data, aes(x = month_date)) +
  # Observed data
  geom_line(aes(y = observed, colour = "Observed"),
    linewidth = 1.2) +
  # Fitted GEE model
  # geom_line(aes(y = fitted, colour = "Fitted"),
  #  linewidth = 1.2) +
  # Counterfactual (pre-PF trend continued)
  geom_line(aes(y = counterfactual,
    colour = "Counterfactual"),
    linewidth = 1.2,
    linetype = "dashed") +
  geom_vline(
    xintercept = as.Date("2024-02-01"),
    linetype = "dotted",
    linewidth = 1
  ) +
  
  scale_colour_manual(
    values = c(
      "Observed" = "black",
      #"Fitted" = "red",
      "Counterfactual" = "blue"
    )
  ) +
  
  labs(
    x = "Month",
    y = "GP medication count",
    colour = "",
    title = "Interrupted Time Series using GEE",
    subtitle = "Dashed blue line shows the expected trend if Pharmacy First had not been introduced"
  ) +
  
  theme_classic(base_size = 14)



```
